---
title: "2_1 Initial data collection report"
author: "Alexander J Malt"
date: "28/11/2019"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```

```{r project, include=FALSE}
library(ProjectTemplate)
load.project()
source("src/initialExploration.R")
```

# Initial data collection report

The University provided dataset from 7 runs of an online cyber-security course. The data provided was in csv format. The data concerned the following areas:

* User activity;
* User personality archetypes;
* User enrollment;
* Leaving survey responses;
* Assessment data;
* Team members; and
* Video statistics.

This data was supported by 7 PDF documents, each providing further information about a particular run of the course (start and end dates, information about each 'step' of the course, etc.).

*N.B. Because this data was provided directly by the University there is no description of the data's background, sources (systems), method of extraction, etc.*