---
title: "Analysing causes of unenrollments"
author: "Alexander J Malt"
date: "28/11/2019"
output:
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```

```{r project, include=FALSE}
library(ProjectTemplate)
load.project()
source("src/Presentation.R")
```


## Aims

- Identify a trend in data relating to unenrollment.
- Identify and assess the reasons students unroll from the course.
- Recommendations in light of data.

## Initial exploration

```{r earlyTrend, echo = FALSE, message=FALSE, warning=FALSE}
initialQuickPlot
```

## Breaking down unenrollment by course run

```{r detailedTrend, echo = FALSE, message=FALSE, warning=FALSE}
unenrollCoursePlot
```

## Explanations?

The pattern seemed to be maintained across

- Age ranges
- Gender
- Employment sector
- Personality archetype
- etc.

## Age ranges

```{r agePlot, echo = FALSE, message=FALSE, warning=FALSE}
agePlot
```


## Gender

```{r genderPlot, echo = FALSE, message=FALSE, warning=FALSE}
genderPlot
```


## New theory: A particular part of the course?

```{r boxPlot, echo = FALSE, message=FALSE, warning=FALSE}
durationPlot
```


## Digging deeper: leaving survey data

```{r leavingData, echo = FALSE, message=FALSE, warning=FALSE}
reasonUnenrolled
```


## Completed weeks of the course

```{r weekCompletion, echo = FALSE, message=FALSE, warning=FALSE}
reasonLeavingBar
```


## Breaking down by course run

```{r reasonLeavingFacet, echo = FALSE, message=FALSE, warning=FALSE}
reasonLeavingFacet
```


## Completed steps

```{r stepCompletion, echo = FALSE, message=FALSE, warning=FALSE}
lastStepPlot
```

## Recommendations

- **Prompts and reminders**: prompt those enrolled to complete bite-sized chunks of the course on a regular basis.
- **Increase quantity of leaving survey data**: offer some money back provided that the leaving survey questionnaire is completed.
